---
title: "figure 1 of Watson and Lovelock"
author: "Helen Wu (Developed based on Robin Hankin's initial scripts)"
date: "07/06/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Functions.R")
```

Note: This is a modified script for a research project concerning Hankin and Mitchell's Daisyworld with Lotka-Volterra Predators.

The original script was developed by Dr. Robin Hankin and can be found at:
https://github.com/RobinHankin/daisyworld

The canonical reference is:

A. J. Watson and J. E. Lovelock 1983. "Biological homeostasis of the
global environment: the parable of Daisyworld".  _Tellus_ 35B:284--289

First, define a function `daisy()` that runs the ODE and returns the
equilibrium configuration, we primarily use this for investigation. 
Parameter values can be changed each run.

```{r definedaisy}
#Function with herbivores 
#Change values within initial_state to investigate differences.
#W (White Daisies), B (Black Daisies), P (Herbivore Predators)

daisy <- function(parameters, ..., all=FALSE){  # something like daisy(params, L=1.2,all=T)
  a <- list(...)
  element <- names(a)
  value <- as.numeric(a)
  parameters[[element]] <- value
  initial_state <- c(W=0.8, B=0.1,P=0.1) # initial population values as %.   
  times <- seq(1,100, by = 1)      # time units for output
  jj <- ode(y = initial_state, times = times, func = watson, parms = parameters)
  if(all){
    return(jj)
  } else {
    out <- c(jj[nrow(jj),-1])  # column 1 of jj is 'time' return every column except first column
  }
  out <- c(value,out)
  names(out) <-c(element,"W","B","P","T_e")
  return(out)
}

```

Define `daisy2()` function, we primarily use this to investigate model outputs 
without Herbivores. This will be used for plotting comparisons below as well.

```{r definedaisy}
#Function without herbivores 
#Change values within initial_state to investigate differences.
#W (White Daisies), B (Black Daisies), P (Herbivore Predators)

daisy2 <- function(parameters, ..., all=FALSE){  # something like daisy(params,L=1.2,all=T)
  a <- list(...)
  element <- names(a)
  value <- as.numeric(a)
  parameters[[element]] <- value
  initial_state <- c(W=0.9, B=0,P=0) # initial population values as %.  
  times <- seq(1,100, by = 1)      # time units for output
  jj <- ode(y = initial_state, times = times, func = watson, parms = parameters)
  if(all){
    return(jj)
  } else {
    out <- c(jj[nrow(jj),-1])  # column 1 of jj is 'time' return every column except first column
  }
  out <- c(value,out)
  names(out) <-c(element,"W","B","P","T_e")
  return(out)
}

```



```{R}
#Store tested results of each run into variable 
#Parameters can be changed in each run, e.g daisy(parameters,L=1.3,all=TRUE)

test<-daisy(parameters,L=1,all=TRUE)   #Model with Herbivores
test2<-daisy2(parameters,L=1,all=TRUE) #Model w/o Herbivores

#Plot test results
plot(test)     #Model with Herbivores
plot(test2)    #Model w/o Herbivores
```

```{r}
#Store results as table

#Convert to dataframe
df <- as.data.frame(test)

DT::datatable(df, 
              extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = TRUE, 
                             paging=TRUE,
                             fixedHeader=FALSE))

```

Now we can run `daisy()` across a range of luminosity `L`:

```{r usedaisy}
#Compute Temperature as a function of Luminosity values and store into matrix
#Calculations for model WITH herbivores

L <- seq(from=0.6,to=2.4,by=0.02) #luminosity
M <- matrix(0,length(L),5)

for(i in seq_along(L)){
  M[i,] <- daisy(parameters,L=L[i])
}
colnames(M) <- c("L","W","B","P","T_e")

head(M)
```

```{r usedaisy}
#Compute Temperature as a function of Luminosity values
#Calculations for model WITHOUT herbivores

L <- seq(from=0.6,to=2.4,by=0.02) #luminosity
M_new <- matrix(0,length(L),5)

for(i in seq_along(L)){
  M_new[i,] <- daisy2(parameters,L=L[i])
}
colnames(M_new) <- c("L","W","B","P","T_e")
head(M_new)

```


```{r}
#Monochromatic Daisyworld Plot
#Plot both with and w/o Herbivores model output in comparison

plot(T_e~L,data=M_new,type='b',main="Planetary temperature as a function of luminosity")
points(T_e~L,data=M,type='l',col='red',lty="dashed")
legend("topleft",col=c("black","red"),legend=c("Monochromatic H & M Daisyworld w/o Herbivores",
                                               "Monochromatic H & M Daisyworld with Herbivores")
,lty=1:2)
```

```{r}
#Bichromatic Daisyworld Plot
#Plot both with and w/o Herbivores model output in comparison
plot(T_e~L,data=M_new,type='b',main="Planetary temperature as a function of luminosity")
points(T_e~L,data=M,type='l',col='red',lty="dashed")
abline(v=0.76, col="cyan3")  #Luminosity = 0.76 (Where oscillations were discovered)
abline(v=1.3, col="cyan3")   #Luminosity = 1.3 (Where oscillations were discovered)
legend("topleft",col=c("black","red","cyan3","cyan3"),
       legend=c("Bichromatic H&M Daisyworld w/o Herbivores",
                "Bichromatic H&M Daisyworld with Herbivores")
,lty=1:2)
text(x=c(0.87,1.4),
     y=c(-5,-5),
     label=c("L=0.76","L=1.3"),
     col=c("cyan4","cyan4"))

```
